Methods

### 2019-10-20 kellen/manuscript.Rmd
Read through methods, waiting for figure access

All GM12878 Hi-C data were downloaded from GSE63525 [@rao20143d] (Supplementary Table1). TAD boundaries were identified using SpectralTAD [@cresswelkg:2019aa]. Three categories of genomic annotations (BroadHMM, histone modifications and transcription factor binding sites (TFBS)) were downloaded from the UCSC genome browser [@Tyner:2017aa] (Supplementary Table 1). All data corresponds to GRCh37/hg19.

@@

We create four types of feature (distance, overlap, percent overlap and combined). Each method starts with binning the matrix based on resolution. For instance, the start of the bins for a 5kb matrix would be located at 0,5kb,10kb,15kb, etc. Distance is calculated by finding the distance between the start of each bin and the closest location of each genomic feature. Overlap is calculated by taking the start point of each bin, flanking it on both sides by half the resolution and calculating the number of times a given genomic feature appears within the flanking region. Percent overlap calculates the percentage of the flanking region that is covered by a given genomic feature. Combined is simply a combination of them all, combined in side-by-side manner. Prediction is performed using matrices of these features with the response being the location of predicted TAD boundaries.

@@

In the contact matrix domains, TAD boundaries are often noisy with imprecise locations. This noise can cause boundaries to shift by multiple bins depending on the TAD caller used. Additionally, we find that genomic features such as CTCF and RAD21 exist at elevated levels in bins around called TAD boundaries and not just in the bin containing the boundary itself. To account for this imprecision we use a flanked form of the precision metric called adjusted precision ($P_a$). In the framework of boundary prediction, traditional precision takes two inputs: true positives ($TP$) whch are correctly identified TAD boundaries and false positives ($FP$) which are regions incorrectly identified as TAD boundaries. Adjusted precision introduces a flanking region ($f$) such that true positives become all predicted TAD boundaries within $f$ of a true TAD boundary and false positives become those that are not within $f$ of a true TAD boundary. We calculate adjusted precision using the following formula:

@@

Conveniently we can convert the adjusted precision score to another measure "adjusted false discovery rate ($FDR_a$)" by simply taking 1-$P_{a}$. $FDR_a$ can be interpreted as the probability that a TAD boundary identified by a given approach is within $f$ of a true TAD boundary.

@@

A feed forward neural network was fit to predict the location of TAD boundaries. The ideal number of hidden layers (1,2) and nodes (4,8,12,16,32,64), and the resampling scheme ("none", "SMOTE", "RUS", "ROS") was tested based on cross-validation using adjusted precision as our measure of accuracy. Results were summarized across all chromosomes. To prevent overfitting, a dropout layer was added between every layer. In general, we find that networks with more than 1 layer tend to either overfit the data (Figure 1), resulting in a decrease in prediction quality. As a result, we present results for single layer models exclusively. Additionally, we find that a dropout of 50% on the final layer provides consistently better results (Figure 2). In addition, we find the choice of activation function to be largely inconsequential to our predictive ability as a result a simple ReLU activation function is used for all layers.

>>>>

All GM12878 Hi-C data were downloaded from GSE63525 [@rao20143d] (Supplementary Table 1 [??? Currently, missing]). TAD boundaries were identified using SpectralTAD [@cresswelkg:2019aa]. Three categories of genomic annotations (BroadHMM, histone modifications and transcription factor binding sites (TFBS)) were downloaded from the UCSC genome browser [@Tyner:2017aa] (Supplementary Table 1). All data corresponds to the GRCh37/hg19 genome assembly.

@@

We create four types of predictors (distance, overlap, percent overlap and combined). Each method starts with binning the matrix based on resolution. For instance, the start of the bins for a 5kb matrix would be located at 0, 5kb, 10kb, 15kb, etc. _Distance_ is calculated by finding the distance between the start of each bin and the closest location [??? TBD, center, or just closest border?] of each genomic feature. _Overlap_ is calculated by taking the start point of each bin, flanking it on both sides by half the resolution and calculating the number of times a given genomic feature appears within the flanking region. _Percent overlap_ calculates the percentage of the flanking region that is covered by a given genomic feature. _Combined_ is simply a combination of them all, combined in side-by-side manner. Prediction is performed using matrices of these features with the response being the location of predicted TAD boundaries.

@@

Given noise and frequent dropouts in Hi-C matrices, especially at high resolution, the location of TAD boundaries detected by a TAD caller is frequently imprecise. The noise can cause boundaries to shift from their true locations by multiple bins depending on the TAD caller used. Additionally, we find that genomic features such as CTCF and RAD21 exist at elevated levels in bins around called TAD boundaries and not just in the bin containing the boundary itself. To account for this imprecision we use a flanked form of the precision metric [??? Need to describe why Precision is used in the first place. Important] called adjusted precision ($P_a$). In the framework of boundary prediction, traditional precision takes two inputs: true positives ($TP$) whch are correctly identified TAD boundaries and false positives ($FP$) which are regions incorrectly identified as TAD boundaries. Adjusted precision introduces a flanking region ($f$) such that true positives become all predicted TAD boundaries within $f$ of a true TAD boundary and false positives become those that are not within $f$ of a true TAD boundary. We calculate adjusted precision using the following formula:

@@

Conveniently we can convert the adjusted precision score to another measure "adjusted false discovery rate ($FDR_a$)" [??? Why we need to? Describe, important] by simply taking $1-P_{a}$. $FDR_a$ can be interpreted as the probability that a TAD boundary identified by a given approach is within $f$ of a true TAD boundary. [??? What is f was set to?]

@@

A feed forward neural network was fit to predict the location of TAD boundaries. The ideal number of hidden layers (1, 2) and nodes (4, 8, 12, 16, 32, 64), and the resampling scheme ("none", "SMOTE", "RUS", "ROS") was tested based on cross-validation [??? Need details, how cross-validation was set up] using adjusted precision as our measure of accuracy. Results were summarized across all chromosomes. To prevent overfitting, a dropout layer was added between every layer. In general, we find that networks with more than 1 layer tend to either overfit the data (Figure 1), resulting in a decrease in prediction quality. As a result, we present results for single layer models exclusively. Additionally, we find that a dropout of 50% on the final layer provides consistently better results (Figure 2). In addition, we find the choice of activation function to be largely inconsequential to our predictive ability as a result a simple ReLU activation function is used for all layers.
